 
---
title: "methyLImp"
author: "Pietro Di Lena"
date: "`r Sys.Date()`"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{methyLImp}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```

```{r setup, include=FALSE}
library(knitr)
library(methyLImp)
```

### The `methyLImp` Package

## Pietro Di Lena

## About the package 
The `methyLImp` implements a missing data imputation method based on single imputation linear regression,
especially designed for and tested on DNA methylation data [1].



## Installation

The package doesn't have any dependencies from other Bioconductor packages.

Installing the latest package from a local copy (assuming it is in the current working
directory of your R session):

```{r eval=FALSE}
install.packages('methyLImp_0.9.9.tar.gz', repos=NULL, type='source')
```

## Trying it out
The package contains a subset of a real 450K Illumina array data, `GSE64495`, which contains beta values
of 100 samples for 200 CpGs with no missing values and it can be used to explore the function quickly:

```{r echo=TRUE}
library('methyLImp')
data(gse64495)
# load in methyLImp dataset
summary(gse64495)
```

## Suggested analysis workflow

# Load data

The methylation data array of either beta or M values has to transposed before imputation,
as variables need to be on the columns and samples on the rows.

# Example workflow

```{r echo=TRUE}

## Load the methyLImp dataset, containing no missing value
data(gse64495)
summary(gse64495)

## Artificially introduce 10% missing values in the first sample
## with the gen_randNA function
set.seed(50)
samp <- 1
frac <- 0.1
gse64495.mis <- gen_randNA(gse64495,samp,frac)
summary(gse64495.mis)

## Impute the missing values with the methyLImp routine.
## Note that variables need to be on the columns and
## samples on the rows.
gse64495.imp <- methyLImp(t(gse64495.mis),min=0,max=0)
gse64495.imp <- t(gse64495.imp)

## Compare imputed and original values
miss <- is.na(gse64495.mis[,samp])
orig <- gse64495[miss,samp]
pred <- gse64495.imp[miss,samp]
gen_stat(orig,pred)

```

```{r echo=FALSE}
sessionInfo()
```

## References

[1] Di Lena P, Sala C, Prodi A, Nardini C. Missing value estimation methods for DNA methylation data. submitted to Bioinformatics




